---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Brian Dinh"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon=false}

## Github Repo Link

[This is the link to the GitHub Repository.](https://github.com/stat301-2-2024-winter/final-project-2-BrianDinh26.git)

:::

```{r}
#| echo: false

# load libraries
library(tidyverse)
library(naniar)
library(here)
```


## Data source

My data source is a used cars dataset from Kirill Lepchenkov on Kaggle.^[[Used Cars Dataset on Kaggle.](https://www.kaggle.com/datasets/lepchenkov/usedcarscatalog/data)]

## Why this data and prediction problem
My prediction objective is to predict the price of cars using data related to used cars online. This will be a regression problem. Having a model would be useful for this objective because it can help potential car buyers assess if they might be overpaying for a used car. Additionally, I am interested in the used car market because I am trying to buy my first car in the summer through the used car market.

## Data quality & complexity check
```{r}
#| echo: true
# data read-in
cars <- read_csv(here("data/cars.csv"))

head(cars, n = 5) |> 
  DT::datatable()
```
The used car dataset has 30 variables. Specifically, there are 10 factor variables, 7 numerical variables, and 13 logical variables. There are 38,531 observations.

### Missingness check
```{r}
#| echo: true

gg_miss_var(cars)
```
Using the `naniar` package, the only variable with missingness is `engine_capacity`, but only 10 instances are missing out of 38,531 total observations, meaning it is not highly significant.

## Potential data issues
```{r}
#| echo: false

cars |>
  ggplot(aes(x = price_usd)) +
  geom_density() +
  labs(x = "Price (USD)",
       y = "Density",
       title = "Density Plot of Price (USD)") +
  theme_classic()
```
Based on the following density plot, it seems that there is a right skew on car prices. There does not appear to be missingness issues. To deal with the right skew, I can transform the dataset using a log transformation. Additionally, I can remove outliers, because the majority of the prices are from the price range of $0 to $15,000.

